Stack Safety as a Security Property

### 2020-06-02 main.tex

\documentclass[conference]{IEEEtran}
\input{macros}

@@

\title{Stack Safety as Security Property}
%\title{The Meaning of/What is Stack Safety}
\iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF
    and...?)}}
\fi

% I get a weird error, so I commented this out for now
%\iftext
%\IEEEauthorblockN{Leonidas Lampropoulos}
%\IEEEauthorblockA{\textit{University of Maryland}}
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address or ORCID
%\and
%\author{\IEEEauthorblockN{Roberto Blanco}
%\IEEEauthorblockA{\textit{\bcp{affiliation?}}
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address or ORCID
%}
%\and
%\IEEEauthorblockN{Benjamin C. Pierce}
%\IEEEauthorblockA{\textit{University of Pennsylvania}
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address or ORCID
%  }
%\and
%\IEEEauthorblockN{Andrew Tolmach}
%\IEEEauthorblockA{\textit{Portland State University}
%% \textit{name of organization (of Aff.)}\\
%% City, Country \\
%% email address or ORCID
%}
%}
%\fi

\maketitle

\iftext
\begin{abstract}
\end{abstract}

\begin{IEEEkeywords}
% component, formatting, style, styling, insert
\end{IEEEkeywords}
\fi

@@

@@

\section{Preliminaries/Background}
%\input{props}

\begin{itemize}
\item machine model
\item Calling conventions
\end{itemize}

\section{Related Work}

\iftext
\section{Introduction}

\begin{itemize}
\item \leo{Not even sure this is true anymore}
\item Noninterference is a security property combining {\em
  confidentiality} and {\em integrity}: secret data should influence
  public outputs and low integrity data should \leo{finish in a pretty
    way}.
\item In this paper we present a stack safety property as a variation
  of noninterference.
\end{itemize}

{\em Hyperproperties}, i.e. properties of pairs\leo{sets?} of
execution traces\footnote{\bcp{I'm not certain we're using the word
    correctly.  E.g., ``A hyperproperty is a set of properties [Clarkson and
    Schneider 2008, 2010]''} \bcp{After more reading, I am still wondering
    if we're using the word the right way.  I think we mean something much
    less general---just a relation on traces.  So I think we should just say
    ``contoured relation'' (and maybe ``lockstep relation'').}}, are the
standard way of expressing security
properties\cn. In particular, {\em noninterference} is commonly
phrased as such a property: given two system states that seem
identical to an observer with low clearance, their execution traces
should also be indistinguishable\cn.

In this paper we focus on a well-behaved closed\bcp{???} subset of
hyperproperties, where the execution traces remain in {\em lockstep}:
the traces execute identical sequences of instructions.
%
This allows us to give a clean extensional characterization of taint
tracking as a noninterference property: intuitively, taint tracking is
a weak lockstep-hyperproperty that guarantees noninterference as long
as the execution traces remain in sync, but then allows for arbitrary
behavior after branching on a secret.
%
We also describe a strong version of lockstep hyperproperties, where
the security policy enforces the lockstepness. \leo{Something about
  full separation}
%
Finally, we generalize strong lockstep hyperproperties to allow for
\leo{what's the correct adjective here? principled? nested?
  restricted?}  changes to the observer clearance, and use this
generalization to characterize a novel stack safety property as
a variation of noninterference.

As it happens, lockstep hyperproperties are also significantly more
amenable to testing compared to traditional hyperproperties. Since
executions remain in lockstep, there is no possibility of \leo{missing
  bugs by never coming back down or a need to find an invariant strong
  enough to prove things}. We have implemented enforcement mechanisms
for all three policies for RISC-V and tested them.

Our contributions are:
\begin{itemize}
\item We present a mathematical abstraction of a class of properties
  called {\em lockstep hyperproperties}, that allow for intuitive
  characterizations of interesting security policies.
\item We give a novel {\em extensional} characterization of taint tracking
  as a form of lockstep-noninterference.
\item We introduce a novel stack safety property that allows for
  the notion of observability to change dynamically throughout
  execution.
\item We develop a testing framework for testing such
  lockstep-hyperproperties, implemented as a parametric co-processor
  for RISC-V.
\end{itemize}

\fi

\section{Preliminaries}

\subsection{Machine model}

Concretely, we have in mind a conventional machine architecture with a
standard (RISC-V) processor extended with some kind of (static or dynamic,
software- or hardware-enforced) security monitor, but we won't need to make
any specific choices about details until we get to \bcp{the implementation
  section}.
% \cref{impl}.

Conventions:
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine components: $\component \in \WORDS \cup \REGS \cup \{
\textsc{pc} \}$
\item Machine states: $\mach \in \MACHS = \COMPONENTS\rightarrow\WORDS$
\end{itemize}

% \bcp{Now we need to define {\em property} formally.  But since we're talking
% about hyperproperties, I'm not certain how to go about it.  The standard way
% would be something like ``A policy $P$ has property $Q$ (i.e., $P$ is
% noninterfering) if, for every pair of traces $\trace{M}$ and $\trace{M'}$
% in $P \times P$, if the initial states $M_0$ and $M'_0$ differ only in
% high parts then every pair ...'' (blah: I'm not sure whether to talk about
% lockstep-corresponding states, which seems too restrictive, or final states,
% which doesn't work if the traces are infinite!  do we really need some kind
% of reactive noninterference here?? Urg.)  (Well, no, we can just do
% a flavor of termination-insensitive NI, which says that all infinite traces
% are indistinguishable from anything...?)}

% \bcp{Should we start from a notion of ``variant traces'' and then talk about
% pairs of variant traces being ``distinguishable'' and then go to the full
% thing?}

The {\em underlying} step function, $\mathord{\stepstounderfull} \in
\MACHS \rightarrow \MACHS$, embodies the standard RISC-V ISA
semantics.
%\bcp{It's probably clearer not to put RISCV in the
%  notation---for most of the paper, the fact that it's a RISCV really
%  doesn't matter.  Moreover, I think we want to make it a partial
%  function, for simplicity of definitions in this part, and if we
%  insist on RISCV here this will confuse people that know it doesn't
%  have a HALT instruction.}

A {\em trace} is a finite or infinite sequence of machine states,
$\seq{\mach} = \mach_0,\, \mach_1,\ldots$, in which adjacent machine
states are related by $\stepstounderfull$---i.e., in which
$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
sequence. We write $\seq{\mach}^i_j$ for the subtrace between $i$ and
$j$.

\subsection{Calling Convention}

\leo{Mainly @Andrew, check if this makes sense for the eager setting.}

We adopt the following calling convention:
\begin{itemize}
\item First the caller increases the stack pointer to allocate stack space
  for arguments and result and zeroes out anything not written.
\item Then a $\JAL$ happens.
\item Then the entry sequence of the callee follows which increases
  the stack pointer again to allocate as much local space as needed
  (e.g. for the return address), and it is all zeroed out if not
  explicitly written. \leo{This can be lifted by dealing with
    stack allocation in the contours}
\end{itemize}

A {\em call map} is a map from a sequence of instructions (read: {\em
  a call sequence}) to a natural number (the number of arguments and
results of the call plus the size of the local state of the callee,
which corresponds to stack locations that will be made available to
the callee).

\subsection{Contours}

A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC, \LOWSEC\} \times \{\HIGHINT, \LOWINT\}$ is a map from
machine components to a pair of confidentiality and integrity labels.
$\LOWSEC$ marks locations whose contents represent data observable
{\em by the current process}; $\HIGHSEC$ corresponds to non-observable
data (e.g. local state of the process caller). $\HIGHINT$ marks
locations that the current process does not have write access to;
$\LOWINT$ is writeable. As examples of different
combinations consider the following:
\begin{itemize}
\item Locations that belong to the current process such as
  available registers, local stack, memory accessible through
  available pointers will be tagged $\LOWSEC \times \LOWINT$:
  they can be both read and written.
\item The local state of the caller that needs to be protected from a
  potentially untrusted callee will be marked $\HIGHSEC \times
  \HIGHINT$. The callee should be able to neither read nor write
  these locations.
\item The instruction memory is read only for every process: $\LOWSEC
  \times \HIGHINT$.
\item Uninitialized memory can be written, but never read. This is only
  applicable in a more lazy setting where we don't zero out memory.
\end{itemize}

\subsection{Identity colorings}
\leo{Feel free to coin another name}

\leo{Went with S for ``self'' and O for ``other''. M/NM was confusing
  with the machine states also being M. T/F didn't really work
  with transitions. Open to other suggestions.}
An {\em identity coloring} of a trace $\seq{\mach}$ is an equal-length
sequence $\seq{\ID}$ of ids ($\IDS = \{\ME, \NOTME, \TRANS\}$, where
$\ID_j = \ME$ if the machine state $\mach_j$ is executing the same
process as $\mach_0$, $ID_j = \NOTME$ if it is executing a different
process, and $ID_j = \TRANS$ during a transition (e.g. during a call
or return sequence).

Given a machine trace $\seq{\mach}$ and a call map the identity coloring can be
computed by iterating through the trace:
\begin{enumerate}
\item Start by coloring $\mach_0$ as $\ME$.
\item Color every subsequent state as $\ME$, until the first machine
  state (if any) at which the instruction executed is a
  the beginning of a call sequence.
\item Color the entire call sequence $\TRANS$. This necessarily
  contains a machine state $\mach_c$ which is a $\JAL$.
\item Color every subsequent state as $\NOTME$, until the first state
  $\mach_r$ after $\mach_c$ in the trace (if any) such that the pc of
  $\PC{\mach_r}$ is equal to $\PC{\mach_c} + 4$, and whose stack
  pointer is identical to that of $\mach_c$. Then mark $\mach_r$ as
  $\ME$ and go to step (2).
\end{enumerate}

\subsection{Contoured Trace}

Given a contour $C$ and a machine state $M$, the {\em contoured trace}
induced by $C$ and $M$, written $\induced{C}{M}$, is the (finite or
infinite) sequence of pairs of (contour, machine state) pairs in which
states in adjacent pairs are related by $\stepstounderfull$, and in which
contours in adjacent pairs are constructed as follows:
\begin{itemize}
\item If in the identity coloring of $\seq{\mach}$ the identity of
  consecutive states is preserved ($\ID_i$ = $\ID_{i+1}$) then
  $C_{i+1} = C_i$.
\item If $\ID_{i} = \ME$ and $\ID_{i+1} = \TRANS$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of $\mach_i$ (the
  caller) to $\HIGHSEC \times \HIGHINT$. The local state is comprised
  of everything below the stack pointer at $\mach_i$.
\item If $\ID_{i} = \TRANS$ and $\ID_{i+1} = \NOTME$, then $C_{i+1}$ is
  constructed from $C_i$ by making the top $n$ locations in the stack,
  where $n$ is the mapping of the transition call sequence in the call map,
  readable and writeable $\LOWSEC \times \LOWINT$.
\item If $\ID_{i} = \TRANS$ \apt {$\NOTME$ ? } and $\ID_{i+1} = \ME$, then $C_{i+1}$ is
  constructed from $C_i$ by setting the local state of the caller
  to $\LOWSEC \times \LOWINT$, and making everything above the stack
  pointer $\HIGHSEC \times \HIGHINT$ again.
\end{itemize}

\subsection{Integrity}

A trace $\seq{\mach}$ satisfies the {\em integrity} safety property
with respect to a contour $\contour$, if for all components
$\component$ where $\contour(\component) = (\_, \HIGHINT)$,
$\mach_i(\component) = \mach_0(\component)$ for all $i$ in the trace.

\leo{For confidentiality we need observations first}

\subsection{Observations}

A {\em component observation} is either a word or the special value
$\bot$.  \leo{Fix notation, clashes with $\LOWINT$}  A {\em state
  observation} is a map from machine components to component
observations.

An {\em observation} is a nonempty, finite or infinite sequence of state
observations in which (1) adjacent elements are always distinct (``no
stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
the the $j^{th}$ state observation in the sequence must also be mapped to
$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
component's value changes from $\bot$ to anything else, it never changes
back to $\bot$ again).

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a contoured trace $\seq{(\contour, \mach)}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour_0(\component) \in \{\LOWSEC \times \_\}$) to its value in
  $\mach_0$ and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where
  $\contour_{i+1}(\component) \in \{\LOWSEC \times \_ \}$) its value
  from $\mach_{i+1}$, and (2) for each hidden component $\component$,
  where $\contour_{i+1} (\component) = \{\HIGHSEC \times \_\}$,
  copying its value from $\stateobs_{i}$, and (3) if the resulting
  $\stateobs_{i+1}$ is identical to $\stateobs_{i}$, then skipping
  $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}
\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

\subsection{Confidentiality}

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$(\LOWSEC, \_)$ by $\contour$---i.e., if $\contour(\component) =
(\LOWSEC, \_)$, then $\mach(\component) = \mach'(\component)$.

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants of each other $N_0
\in \variant{C}{M_0}$, then the observations of one is a prefix of the
observations of the other: $\obstrace(\seq{\contour, \mach}) \subseteq
\obstrace(\seq{\contour, N})$.

\apt{Does this differ from lockstep equivalence? And why prefix?}

\subsection{Stack Safety}

A trace $\seq{\mach}$ satisfies {\em stack safety} with respect to a
contour $\contour$ if:
\begin{itemize}
\item $\seq{\mach}$ satisfies integrity with respect to $\contour$.
\item For every variant $N \in \variant{C}{\mach_0}$, $\seq{\mach}$
  and the the induced trace $\seq{N}$ satisfy confidentiality with
  respect to $\contour$.
\item Every subtrace $\seq{\mach}^r_c$ colored $\NOTME$ in the
  identity coloring of $\seq{\mach}$ satisfies stack safety with
  respect to the contour $\contour_c$.
\end{itemize}

\bcp{Needs to be formulated coinductively?!}

\bcp{What about unitialized memory (written by previous calls)}

\section{Related work}

>>>>

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\input{macros}

@@

%% Title information
\title{Stack Safety as a Security Property}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
%\author{First1 Last1}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
%  \institution{Institution1}            %% \institution is required
%  \streetaddress{Street1 Address1}
%  \city{City1}
%  \state{State1}
%  \postcode{Post-Code1}
%  \country{Country1}                    %% \country is recommended
%}
%\email{first1.last1@inst1.edu}          %% \email is recommended
% 
%%% Author with two affiliations and emails.
%\author{First2 Last2}
%\authornote{with author2 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%\affiliation{
%  \position{Position2a}
%  \department{Department2a}             %% \department is recommended
%  \institution{Institution2a}           %% \institution is required
%  \streetaddress{Street2a Address2a}
%  \city{City2a}
%  \state{State2a}
%  \postcode{Post-Code2a}
%  \country{Country2a}                   %% \country is recommended
%}
%\email{first2.last2@inst2a.com}         %% \email is recommended
%\affiliation{
%  \position{Position2b}
%  \department{Department2b}             %% \department is recommended
%  \institution{Institution2b}           %% \institution is required
%  \streetaddress{Street3b Address2b}
%  \city{City2b}
%  \state{State2b}
%  \postcode{Post-Code2b}
%  \country{Country2b}                   %% \country is recommended
%}
%\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~General programming languages}
%\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{Stack Safety, Micropolicy, Random Testing, RISC-V}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

@@

\leo{Variations of these should probably go here:}
    {\it Eager stack integrity} states that if a function call returns, the caller's stack data
    will be identical to what it was before the call. This implies the weaker {\it observable
    stack integrity}, which states that any changes to the caller's frame will produce the same
    observable behavior.

    {\it Eager stack confidentiality} states that data outside a callee's stack frame is secret,
    and does not interfere with its observable behavior or with the machine state at return. Its
    weaker variant {\it observable stack safety} lifts the latter restriction and states that
    the callee may not leak these secrets directly or by changing the behavior of its caller.

@@

\section{Background}

We begin by describing our machine model. Concretely, we have in mind a
conventional machine architecture with a standard (RISC-V) processor
extended with some kind of (static or dynamic, software- or
hardware-enforced) security monitor, but we won't need to make any
specific choices about details until we get to \bcp{the policy one, probably}.
%In this section we will describe an abstract interface to our machine model.

Values and addresses are words drawn from some set of all words $\WORDS$. We will
denote words with $\word$.\leo{TODO: If we ever need ``A'' come back to this.}

Components are words (denoting addresses) and registers.
%
    \[\reg \in \REGS\]
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The registers must include two special purpose registers: the program
counter \PCname and the stack pointer \SP.

A machine state is just a map from components to values:
%
\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]

Our stack safety property will be defined in terms of {\em observations}, which may be
silent (written $\tau$) or a word of data:
%
\[\obs \in \OBSS = \tau ~ | ~ \word \in \WORDS\]

These observations are a byproduct of the step function of the
machine, which takes a machine state and produces a pair of a
machine state and an observation.
%
\(\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \).

\paragraph*{Policy}

A policy is a general model of an enforcement mechanism, consisting of
a set of policy states, \(\pol \in \POLS\) and a policy step function
\((\mach, \pol) \stepstopol \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\).

%    The initializer and step function are both partial functions,
%    the former to reflect static guarantees (some initial states may not be valid at all) and
%    the latter to reflect dynamic fail-stop behavior. We then define machine-policy pairs and
%    a step function on them:

Let $\MPS$ be the set of all pairs of machine states and policy states.
%
\[\MPS = \MACHS \times \POLS\]
%
Then we can lift this policy step function to operate on such pairs,
by combining it with the regular step function for machine states.

%    \[\stepstoobs{}_P \subseteq \MPS \times \MPS\]

    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} (\mach_1, \pol_1) \stepstopol \pol_2}
           {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]
\leo{Notation: Should $\stepstoobs{}$ be partial?}

\paragraph{Traces}

A trace over some type $A$ is a potentially infinite, nonempty
sequence of $A$s. We will mostly be concerned with sequences of
machines states, usually denoted $\machT$, and with sequences of
triples of (machine states, policy states, and observations), usually
denoted $\MPT$.

\leo{I tried the ? notation for the potentially empty trace, which reveals that
  we're slightly informal in the second case. Expand?}
We define the append operation \(T \cdot T?\), where \(T?\) may be a trace or empty (\(\varepsilon\)):
\[T \cdot T? \triangleq
  \begin{cases}
    T & \text{if } T? = \varepsilon \\
    t T? & \text{if } T = t \\
    t T' \cdot T? & \text{if } T = t T' \\
  \end{cases}\]

We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
%
\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
       {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
%
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

We will also wish to take a prefix of the trace from an initial state
up through the first state where some condition holds. Let \(f\) be
such a predicate on machine states.  Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix from
\(\mach\) until \(f\)''):\leo{Abrupt. Explain in words?}
    \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \neg f\ \mach_1 \hspace{.5cm} \mach_1
              \hookrightarrow \machT | f}
           {\mach_0 \hookrightarrow (\mach_0,\obs) \machT | f}\]
    \[\frac{f\ m}
           {m \hookrightarrow (m,\tau)}\]

Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
state pair to the trace of triples of machine states, policy states,
and observations induced by by \(\stepstoobs{}_P\):
%
\leo{I think this would read better as not exists. Thoughts?}
\[\frac{\forall \pol_1 . \pol_0 \not \longrightarrow \pol_1}
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \longrightarrow \pol_1 \hspace{.5cm}
  \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]

We also lift the prefix-up-to relation to such traces.
\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm}
  \neg f\ \mach_1 \hspace{.5cm} \mpstate{\mach_1}{\pol_1} \hookrightarrow_P \machT | f }
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT | f}\]

\[\frac{\mpstate{\mach_0}{\pol_0} \xrightarrow{\obs} \mpstate{\mach_1}{\pol_1} \hspace{.5cm} f\ \mach_1}
       {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs}
         \mpostate{\mach_1}{\pol_1}{\tau} | f}\]

Finally, we will use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations.

%    \COQ{this operation steps until it reaches a state on which \(f\) holds,
%    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
%    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
%    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
%      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
%    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}

\paragraph*{Observations and Equivalence}

We say that a trace of observations $\obsT_1$ is a prefix of $\obsT_2$
\leo{Benjamin suggested ``as far as an attacker is conerned in an
  e-mail, but we haven't said anything about attackers yet.}, if the
sequence of word-observations of $\obsT_1$ is a prefix of those of
$\obsT_2$, that is we operate up to deletion of \(\tau\) observations, coinductively:

\[\frac{} {\obsT \lesssim \obsT}\]

\[\frac{} {\tau \lesssim \obsT}\]

\[\frac{\obsT_1 = \tau \obsT_1' \hspace{.5cm} \obsT_1' \lesssim \obsT_2}
       {\obsT_1 \lesssim \obsT_2}\]

\[\frac{\obsT_2 = \tau \obsT_2' \hspace{.5cm} \obsT_1 \lesssim \obsT_2'}
       {\obsT_1 \lesssim \obsT_2}\]

\[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
       {\obsT_1 \lesssim \obsT_2}\]

% Fixed by Andrew:
%\COQ{This is the same as the coq version minus a redundant case.}

We then define equivalence of observation traces as traces prefixing each other:
\[O_1 \simeq O_2 \triangleq O_1 \lesssim O_2 \land O_2 \lesssim O_1\]

\leo{Should we point out here the edge cases of all-tau traces?}

% Fixed by Andrew:
%    \COQ{this should be equivalent to the coq version, and we should adopt it.}
%    \leo{We had this, but it was much harder to work with for my proofs. We should still
%      prove them equivalent if we want to present it this way (which is probably cleaner for
%      presentation purposes)}

Let the predicate \(\isinf(T)\) hold when the trace $T$ is infinite.
We then define {\em policy-sensitive observational equivalence}: a
machine-policy-observation\leo{Call these MPO traces?} is equivalent
to a machine-observation trace, if the MPO trace is infinite and the
observations of MPO and M coincide, or if the MPO trace is finite
(e.g. due to a policy fault) and its observations are a prefix of
those of M.

\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \lesssim \mathit{Obs}(\machT)}
           {\MPT \sim_P \machT}\]

\[\frac{\isinf(\MPT) \hspace{.5cm} \mathit{Obs}(\MPT) \simeq \mathit{Obs}(\machT)}
           {\MPT \sim_P \machT}\]

\section{Stack Safety, Formally}

In this section we will describe stack safety as a formal
property.
\leo{Should call maps and returns go here or in preliminaries?}

\paragraph*{Call Maps and Returns}

A {\it call map} $\callmap$ identifies machine states that represent a
call, abstracting away from the calling convention of any concrete
machine. Formally $\callmap$ is a partial map from machine states to a
natural number, the number of arguments of that call.
%
\[\callmap \in \CALLMAPS = \MACHS \rightharpoonup \mathbb{N}\]

For concreteness, in a typical calling convention, a call map may be
considered to identify the state just before the \(\JAL\), in which
space for arguments is already allocated, but the mechanism is more
general.

A state on which a call map is defined is referred to as a {\it call
  state}, and we consider the call to have returned the first time
control returns to the instruction following the \(\JAL\) with the
stack restored. Formally this is a relation between the states:

\[\begin{split}
\ret{\mach_c}\ \mach_r \triangleq & \mach_r(\PCname) = \nextPC(\mach_c(\PCname)) \land \\
& \mach_r(\SP) = \mach_c(\SP)
\end{split}\]

\leo{How do we initialize the first policy state based on a call map? It's not
  part of the policy. Do we parameterize policies?}

\paragraph*{Stack safety}

A stack safety property is a property of a transition system, a
policy, and a callmap that is decomposed into an integrity property and a
confidentiality property.

    \[\begin{split}
      \textit{stack safety} \triangleq \textit{stack integrity} \land \textit{stack confidentiality} \\
    \end{split}\]

\leo{TODO: Rephrase from introduction} Informally, {\it Stack
  integrity} states that if a function call returns, the caller's
private stack data will be identical to what it was before the
call. {\it Stack confidentiality} states that data outside a callee's
stack frame is secret, and do not interfere with its observable
behavior or with the machine state at return.

As is standard with security properties, we associate with each
component of the machine a {\em label}: a pair of security levels
denoting high or low confidentiality and high or low integrity.
%
\[\mathit{label} ::= \{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT,\LOWINT\}\]
%
Since we need to enforce integrity and confidentiality at {\em every}
function call in a trace, we must treat the security level of data
dynamically. To that end, we introduce the notion of a {\em contour},
a map from components to labels:
%
\[\contour \in \CONTOURS ::= \COMPONENTS \rightarrow \mathit{label}\]

\leo{What about instructions? Shouldn't they be LC/HI?}
When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\)
for some \(n\), we can construct a contour based on the bounds of the
callee's stack frame, as follows:
\leo{Should the first two cases have a $k \not \in \REGS$?}
\[\mathit{Cof}(\mach,n)(\component) =
\begin{cases}
  (\HIGHSEC,\HIGHINT) & \text{if } \component \leq \mach(\SP) - n \\
  (\HIGHSEC,\LOWINT) & \text{if } \component \geq \mach(\SP) \\
  (\LOWSEC,\LOWINT) & \text{else} \\
\end{cases}\]
%
The private state of the callee is the entire area of the memory below
the stack pointer, with the exception of the $n$ locations
corresponding to the arguments of a call. These components are tagged
high confidentiality and high integrity: the callee should be able to
neither read, nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
Registers and the frame of arguments to the callee are free for the
callee to use, and are therefore tagged with low integrity and low
confidentiality.

\leo{Shouldn't this be a definition instead of a paragraph?}
%\paragraph{Integrity}

\definition{Stack Integrity}

A system satisfies integrity with respect to a callmap \(\callmap\) if
for any initial state \(\mach_0\), call state \(\mach_c\) where
\(\callmap(\mach_c) = n\), and policy state \(\pol_c\), and return
state \(\mpstate{\mach_r,\pol_r}\), where: \leo{words: induced trace, call in the trace, take the prefix}
%
\[\mach_0,\mathit{pinit}(\mach_0) \hookrightarrow ... \mpstate{\mach_c}{\pol_c,\_} ... \mpstate{\mach_r}{\pol_r,\_} | \ret{\mach_c}\]
%
it is always the case that: \leo{words here too: the value of confidential components is unchanged}
\[\forall \component . \contour(\component) = \HIGHINT \rightarrow \mach_c(\component) = \mach_r(\component)\]

\definition{Stack Confidentiality}
